### 2025-05-16 thesis_julio/Chapter1/chapter1.tex
commit5
may16

\begin{figure}[h]
    \centering
    \includegraphics[scale=1]{table1.png}
    \caption{The magnitude of the four forces compared}
    \label{fig:four-forces}
\end{figure}

@@

\includegraphics[scale=0.7]{sm.png}

@@

 \includegraphics[scale=0.35]{LHC.png}

@@

The performance of a collider is determined by the beam energy and the luminosity, the luminosity is a key parameter in particle colliders is a quantity that measures the ability of a particle accelerator to produce a required number of interactions and is given by: \cite{Lum}

$\frac{dR}{dt} = \mathcal{L} \cdot \sigma_{p} $

@@

 \includegraphics[scale=0.80]{lumi.png}

@@

$ 2N_{1}N{2}fN_{b} \int \int \int^{\infty}_{-\infty}  \rho_{1x} (x) \rho_{1y}(y) \rho_{1s} (s-s_{0}) \rho_{2x}(x) \rho_{2y}(y) \rho_{2s} (s+s_{0}) dxdydsds_{0}  $

@@

$ L_{int} = \int L (t') dt' $

@@

$\mathcal{L}_{int} \cdot \sigma_{p} =$ number of events of interest

@@

 \includegraphics[scale=1.25]{integratedlum.png}

>>>>

\begin{figure}[h]
    \centering
    \includegraphics[scale=1]{table1.png}
    \caption{Comparison of the strength magnitude of the four fundamental forces}
    \label{fig:four-forces}
\end{figure}

@@

\begin{figure}[h]
    \centering
   \includegraphics[width=1\textwidth]{sm.png}
    \caption{The standard model of particle physics}
    \label{fig:four-forces}
\end{figure}

@@

\begin{figure}[h]
    \centering
   \includegraphics[width=1\textwidth]{LHC.png}
    \caption{The accelerator facility at cern for the LHC}
    \label{fig:LHC}
\end{figure}

@@

%The performance of a collider is determined by the beam energy and the luminosity, in order to compute the luminosity we need to define the cross section of the accelerator first this is for the calculation of the interest rates, which have to take in consideration the flux which is defined as the number of particles crossing a unit are per unit of time, the cross section is denoted by $\sigma$ and every scattering process has a cross section $\simga_{i}$ but the one that we are interested on is the total cross section which is the sum of all the $\simga_{i}$.

The performance of a collider is determined by the beam energy and the luminosity. The cross section It's known as parameter like "the size of a target"

%The define of cross section can be understand in a situation where a single particle is travelling to a velocity $v _{a}$in a region defined by A, which contains $n_{b}$ particles of a particle type b moving to a velocity $v_{b}$ in the opposite direction of $v_{a}$ if in a time $\delta t$  the particle a cross a region contained a total of $nb(v_{a} + v_{b})A\delta t$ particles of type b, the interaction probability can be obtained from the effective total cross seciotanl area divided by A which can be seen as the probability of that hte incident particles passes through one of the region of area $\sigma$ around each of the $\delta N$ particles then we can talk about the total rate being:

\begin{equation}
rate = flux x number of target particles x cross section
\end{equation}

And we can define the cross section as follow:

\begin{equation}
\sigma = \frac{number of interaction per unit of time per target particle}{incident flux}
\end{equation}

The luminosity is a key parameter in particle colliders is a quantity that measures the ability of a particle accelerator to produce a required number of interactions and is given by: \cite{Lum}

\begin{equation}
\frac{dR}{dt} = \mathcal{L} \cdot \sigma_{p} 
\end{equation}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{lumi.png}
    \caption{Two beams colliding}
    \label{fig:beamslumi}
\end{figure}

@@

\begin{equation}
 2N_{1}N{2}fN_{b} \int \int \int^{\infty}_{-\infty}  \rho_{1x} (x) \rho_{1y}(y) \rho_{1s} (s-s_{0}) \rho_{2x}(x) \rho_{2y}(y) \rho_{2s} (s+s_{0}) dxdydsds_{0}  
\end{equation}

@@

\begin{equation}
 L_{int} = \int L (t') dt' 
\end{equation}

@@

\begin{equation}
L_{int} \cdot \sigma_{p} = number of events of interest 
\end{equation}

@@

\begin{figure}[h]
    \centering
     \includegraphics[scale=1.25]{integratedlum.png}
     \caption{Integrated luminosity at CMS}
    \label{fig:CMS-Luminosity}
\end{figure}

### 2025-05-16 thesis_julio/Chapter2/chapter2.tex
may16

The Compact Muon Solenoid (CMS) detector is one of the four interactions points on the LHC. It consist on different detectors made for different purposes, the silicon tracker, the electromagnetic calorimeter (ECAL), the hadron calorimeter (HCAL) the superconducting solenoid and the muon chamber. 
\\
\begin{figure}[h]
    \centering
     \includegraphics[width=1\textwidth]{cms2.png}
     \caption{The CMS detector size comparison with each of their components}
    \label{fig:four-forces}
\end{figure}

\\
The CMS has the form of a cylindrical onion with several concentric layers of components, this components are important to determine the properties of different particles all of this by bending their trajectories with a magnet that bend charged particles, this helps to identify the charges of the particles since they bend in different directions and allows to measure the momentum of a particle sine high momentum particles bend less than the low momentum ones. The principal magnet of the CMS is called Super Conducting Solenoid this magnet generates a field of about 4 Teslas and is very large 6.3m cold bore, 12.5 meters of length and 200 t of mass. The high magnetic field is confined to the volume of the detectors using the "return yoke", this yoke have a magnetic field strength of 2 Teslas and acts as the main support as well as the muon filter Its made of a high amount of steel and It weights more than 11000 tonnes.     \cite{CMS3}

@@

\begin{figure}[h]
    \centering
     \includegraphics[width=1\textwidth]{cms3.png}
     \caption{Layout of the CMS Innertracker includin the Pixel Detector, and other substructures TIB, TOB, TID and TEC}
    \label{fig:inner-track}
\end{figure}

@@

Having information of the energy is important to understand what is happening at the collision point for this purpose there are two calorimeters on the CMS. The ECAL is made of 61200 lead tungstate crystals in the central barrel cap closed by 7324 crystals in the two of the endcaps in front of this crystals there is a press shower detector, the high density crystals allow  the calorimeter to be fast, have a fine granularity and being radiation resistant, Its capabilities are improved by the good crystal resolution which are provided by a homogeneous crystal calorimeter. It's surrounding the tracker and aims to measure the the energy of electrons, positrons or photons and stop them completely. The other calorimeter is the HCAL which is surrounding the ECAL and plays an essential role on the measurements of the energy of hadrons like the quarks, gluons, neutrinos and some exotic particles. which are also stopped here.  
\\

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{cms.png}
     \caption{Different particles travelling across the different detectors of the CMS experiment}
    \label{fig:four-forces}
\end{figure}

\\
The CMS was optimized for the detections of muons and is the last particle that the CMS observe directly with the help of the muon chambers that is composed by many subdetectors which are interleaved with the return of the yoke of the solenoid located after the super conducting solenoid, muons are similar to electrons but 200 times heavier so they are not stopped by the calorimeters, the muon chambers have 3 main purposes; muon identification, momentum measurement and triggering. \cite{CMS2}

@@

For the tracking

\section{Pixel Cluster Counting}

The Pixel Cluster Counting (PCC) method consist in counting the average number of pixel clusters on the detector to measure luminosity. This occurs during a zero bias event, an event that is trigged by requiring only that two proton bunches cross at the CMS interaction point. Giving that the number of pixels in really big the probability that a pixel is being hit by two different tracks by the same bunch crossing is really small. The mean number of pixel clusters in a simulated zero bias event is in the order of 100 per pp collision. The number of pixel clusters per bunch crossing is linearly dependent on pileup and therefore an accurate measure of the instantaneous luminosity.\cite{PCC1} The mean number of pixel cluster per trigger is \\

@@

here f is the frequency of the LHC, defining a new quantity $\simga_{cluster}$ = $\simga_{i}<N_{Pixel/Interaction}> $ and combining equation (2.1) and (2.2) we got

@@

From the total of 1856 modules on the pixel detector (1184 on the BPIX and 672 on FPIX) so in the analysis for the pcc vdm calibration onf 2024 for the fill 9639a total of 240 modules were used.

>>>>

The Compact Muon Solenoid (CMS) detector is one of the four interactions points on the LHC. It consist on different detectors made for different purposes, the silicon tracker, the electromagnetic calorimeter (ECAL), the hadron calorimeter (HCAL) the superconducting solenoid and the muon chamber.

\begin{figure}[H]
    \centering
     \includegraphics[height=0.73\textwidth, width=1\textwidth]{cms2.png}
     \caption{The CMS detector size comparison with each of their components}
    \label{fig:four-forces}
\end{figure}

The CMS has the form of a cylindrical onion with several concentric layers of components, this components are important to determine the properties of different particles all of this by bending their trajectories with a magnet that bend charged particles, this helps to identify the charges of the particles since they bend in different directions and allows to measure the momentum of a particle sine high momentum particles bend less than the low momentum ones. The principal magnet of the CMS is called Super Conducting Solenoid this magnet generates a field of about 4 Teslas and is very large 6.3m cold bore, 12.5 meters of length and 200 t of mass. The high magnetic field is confined to the volume of the detectors using the "return yoke", this yoke have a magnetic field strength of 2 Teslas and acts as the main support as well as the muon filter Its made of a high amount of steel and It weights more than 11000 tonnes.     \cite{CMS3}

@@

%\begin{figure}[H]
   % \centering
     %\includegraphics[width=1\textwidth]{cms3.png}
     %\caption{Layout of the CMS Innertracker includin the Pixel Detector, and other substructures TIB, TOB, TID and TEC}
    %\label{fig:inner-track}
%\end{figure}

@@

Having information of the energy is important to understand what is happening at the collision point for this purpose there are two calorimeters on the CMS. The ECAL is made of 61200 lead tungstate crystals in the central barrel cap closed by 7324 crystals in the two of the endcaps in front of this crystals there is a press shower detector, the high density crystals allow  the calorimeter to be fast, have a fine granularity and being radiation resistant, Its capabilities are improved by the good crystal resolution which are provided by a homogeneous crystal calorimeter. It's surrounding the tracker and aims to measure the the energy of electrons, positrons or photons and stop them completely. The other calorimeter is the HCAL which is surrounding the ECAL and plays an essential role on the measurements of the energy of hadrons like the quarks, gluons, neutrinos and some exotic particles. which are also stopped here.

\begin{figure}[H]
    \centering
    \includegraphics[width=1\textwidth]{cms.png}
     \caption{Different particles travelling across the different detectors of the CMS experiment}
    \label{fig:four-forces}
\end{figure}

The CMS was optimized for the detections of muons and is the last particle that the CMS observe directly with the help of the muon chambers that is composed by many subdetectors which are interleaved with the return of the yoke of the solenoid located after the super conducting solenoid, muons are similar to electrons but 200 times heavier so they are not stopped by the calorimeters, the muon chambers have 3 main purposes; muon identification, momentum measurement and triggering. \cite{CMS2}

@@

\section{Cluster Identification}

The Pixel Cluster Counting (PCC) method consist in counting the average number of pixel clusters on the detector to measure luminosity. This occurs during a zero bias event, an event that is trigged by requiring only that two proton bunches cross at the CMS interaction point. Giving that the number of pixels in really big the probability that a pixel is being hit by two different tracks by the same bunch crossing is really small. The mean number of pixel clusters in a simulated zero bias event is in the order of 100 per pp collision. The number of pixel clusters per bunch crossing is linearly dependent on pileup and therefore an accurate measure of the instantaneous luminosity.\cite{PCC1} The mean number of pixel cluster per trigger is

\begin{figure}[h]
    \centering
    \includegraphics[width=1\textwidth]{cluster2.jpg}
    \caption{Cluster}
    \label{fig:cluster}
\end{figure}

@@

here f is the frequency of the LHC, defining a new quantity $\sigma_{cluster}$ = $\sigma_{i}<N_{Pixel/Interaction}> $ and combining equation (2.1) and (2.2) we got

@@

A module veto list is created that removes modules that have unstable behavior in the cluster counting, to the determine the number of modules excluded the stability of their contribution to the total (the module weight) is studied. This selection is made using the zero bias datasets using two steps, first a selection on the stability over time and then a selection of stability vs instantaneous luminosity. For the stability selection different datasets are used, depending on the dataset you obtain time profiles per module and the absolute value of the deviation from the mean weight and averaged, this process is iterate for every module to remove outlier modules. For the linearity selection the module weight is evaluated as a function of quantity which is proportional to the instantaneous luminosity this quantity being the average PCC per module, the image (2.6) shows the difference between a good module and a bad module.

\begin{figure}[H]
    \centering
    \includegraphics[height=0.35\textwidth ,width=1\textwidth]{module.png}
    \caption{Comparison between two modules with average PCC in 100 LS. Good module on the left, bad module on the right.}
    \label{fig:modules}
\end{figure}

To obtain the number of modules that are damaged and can exhibit issues with stability and linearity and can affect the final luminosity measurements the following steps were followed:

\begin{itemize}
  \item First, all of the BPIX Layer 0 Modules are removed since they are consistently unstable.
  \item Then modules based on the noise level are removed in the vdm fill super separation data. 
  \item Remove modules with no clusters in the vdm fill this means with weight=0 
  \item Compare stability parameters between modules and remove outliers in the vdm fill
  \item Compare stability parameter for each block of data and perform a second iteration if necessary
  \item Compare linearity parameter between modules and remove outliers
\end{itemize}

The pixel detector has a total of 1856 modules (1184 on the BPIX and 672 on FPIX) in this analysis for the pcc vdm calibration on 2024 for the fill 9639, a total of 1616 modules were removed for a total of 240 modules with 193 in the BPIX part of the detector and the other ones on the FPIX part.

### 2025-05-16 thesis_julio/Portada/Portada.tex
commit1
may16

 %% Aquí comienza el documento.
\thispagestyle{empty} %% PARA QUITAR FORMATO DE PIES Y ENCABEZADOS DE PÁGINA A LA PORTADA
%
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%% MINIPAGE PARA EL LOGO DEL LADO IZQUIERDO (UNIVERSIDAD) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%\fbox{
\begin{minipage}[c][0.15\textheight][c]{0.25\textwidth}
    \begin{center}
        \includegraphics[height=3.2cm, keepaspectratio=true]{unison-logo.png} %% LOGO IZQUIERDO
    \end{center}
\end{minipage}
%}
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%%%%%% MINIPAGE PARA EL CUERPO DE LA PORTADA (NOMBRES, TITULOS, ETC.) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%
%
%
%
%\fbox{
\begin{minipage}[c][0.17\textheight][t]{0.7\textwidth}
    \begin{center}
        {\scshape\LARGE UNIVERSIDAD DE SONORA} %% NOMBRE DE LA UNIVERSIDAD
        \vspace{.5cm}   %% DEFINE EL ESPACIO ENTRE EL NOMBRE DE LA UNIVERSIDAD Y LA LÍNEA DOBLE
        \hrule height2.5pt  %% DEFINE EL ANCHO DE LA PRIMER LÍNEA
        \vspace{.1cm}    %% DEFINE EL ESPACIO ENTRE LAS DOS LINEAS
        \hrule height1pt  %% DEFINE EL ANCHO DE LA SEGUNDA LINEA (NORMALMENTE ES MAS DELGADA)
        \vspace{.4cm}   %% DEFINE EL ESPACIO ENTRE LA DOBLE LINEA Y EL NOMBRE DE LA DIVISIÓN
        {\scshape \large Facultad Interdisciplinaria de Ciencias Exactas y Naturales}
        {\scshape Departamento de Investigaci\'on en F\'isica} %% NOMBRE DE LA DIVISIÓN
    \end{center}
\end{minipage}
%}
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%%%%%%%%%%%%%%%% MINIPAGE PARA EL LOGO DEL LADO DERECHO (DIVISIÓN ACADÉMICA)
%
%
%
%
%\fbox{

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%
%  ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%
%%%%%%%%%%%%%%%% 5TO MINIPAGE PARA EL CUERPO DE LA PORTADA (NOMBRES, TÍTULO, ETC.) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%\fbox{
\begin{minipage}[l][0.78\textheight][t]{0.1\textwidth}
    \begin{center}
    \hskip0pt
    \vrule width2.5pt height17cm    %% height definde la altura de la linea gruesa
        \hskip1mm
        \vrule width1pt height17cm  %% height define la altura de la linea delgada
        \end{center}
\end{minipage}
%}
%\fbox{
\begin{minipage}[c][0.78\textheight][t]{0.8\textwidth}
      \begin{center}
      \vspace{2cm}
        {\Large \scshape {Development and calibration of the luminosity measurement system CMS Silicon Pixel Detector}}

@@

        \makebox[5cm][c]{\LARGE THESIS}  \\[20pt]
         in partial fulfillment of the
requirements for the degree of:\\[5pt]
        {\Large \textbf{{Maestr\'ia en Ciencias (F\'isica)}}}\\[40pt]
        By:\\[12pt]
        \textbf{  {\large Luis Enrique CUEVAS PICOS}}

@@

        { \large Director}:\\Dr. Jose Feliciano BENITEZ RUBIO

@@

        { Hermosillo, Sonora} \hskip2.5cm {August 15, 2023}
      \end{center}
\end{minipage}
%}

>>>>

 %% Aquí comienza el documento.
\thispagestyle{empty} %% PARA QUITAR FORMATO DE PIES Y ENCABEZADOS DE PÁGINA A LA PORTADA
%
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%% MINIPAGE PARA EL LOGO DEL LADO IZQUIERDO (UNIVERSIDAD) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%\fbox{
\begin{minipage}[c][0.15\textheight][c]{0.25\textwidth}
    \begin{center}
        \includegraphics[height=3.2cm, keepaspectratio=true]{unison-logo.png} %% LOGO IZQUIERDO
    \end{center}
\end{minipage}
%}
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%%%%%% MINIPAGE PARA EL CUERPO DE LA PORTADA (NOMBRES, TITULOS, ETC.) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
%
%
%
%
%\fbox{
\begin{minipage}[c][0.17\textheight][t]{0.7\textwidth}
    \begin{center}
        {\scshape\LARGE UNIVERSIDAD DE SONORA} %% NOMBRE DE LA UNIVERSIDAD
        \vspace{.5cm}   %% DEFINE EL ESPACIO ENTRE EL NOMBRE DE LA UNIVERSIDAD Y LA LÍNEA DOBLE
        \hrule height2.5pt  %% DEFINE EL ANCHO DE LA PRIMER LÍNEA
        \vspace{.1cm}    %% DEFINE EL ESPACIO ENTRE LAS DOS LINEAS
        \hrule height1pt  %% DEFINE EL ANCHO DE LA SEGUNDA LINEA (NORMALMENTE ES MAS DELGADA)
        \vspace{.4cm}   %% DEFINE EL ESPACIO ENTRE LA DOBLE LINEA Y EL NOMBRE DE LA DIVISIÓN
        {\scshape \large Facultad Interdisciplinaria de Ciencias Exactas y Naturales} \\
        {\scshape Departamento de Investigaci\'on en F\'isica} %% NOMBRE DE LA DIVISIÓN
    \end{center}
\end{minipage}
%}
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%
%   ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%%%%%%%%%%%%%%%%%%%%%%% MINIPAGE PARA EL LOGO DEL LADO DERECHO (DIVISIÓN ACADÉMICA)
%
%
%
%
%\fbox{

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%
%  ESPACIO COMENTADO PARA EVITAR DESAJUSTES POR GENERACIÓN DE NUEVO PÁRRAFO (NO DEJAR ESPACIOS ENTRE MINIPAGES)
%
%
%
%
%
%
%%%%%%%%%%%%%%%% 5TO MINIPAGE PARA EL CUERPO DE LA PORTADA (NOMBRES, TÍTULO, ETC.) %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%
%\fbox{
\begin{minipage}[l][0.78\textheight][t]{0.1\textwidth}
    \begin{center}
    \hskip0pt
    \vrule width2.5pt height17cm    %% height definde la altura de la linea gruesa
        \hskip1mm
        \vrule width1pt height17cm  %% height define la altura de la linea delgada
        \end{center}
\end{minipage}
%}
%\fbox{
\begin{minipage}[c][0.78\textheight][t]{0.8\textwidth}
      \begin{center}
      \vspace{2cm}
        {\Large \scshape {Calibration of the Pixel Detector at CMS}}

@@

        \makebox[5cm][c]{\LARGE THESIS}  \\[20pt]
         in partial fulfillment of the
requirements for the degree of:\\[5pt]
        {\Large \textbf{{Maestr\'ia en Ciencias (F\'isica)}}}\\[40pt]
        By:\\[12pt]
        \textbf{  {\large JULIO C\'ESAR BORB\'ON FRAGOSO}}

@@

        { \large Director}:\\DR. JOSE FELICIANO BENITEZ RUBIO

@@

        { Hermosillo, Sonora} \hskip2.5cm {2025}
      \end{center}
\end{minipage}
%}
